---
title: "Bioinformatics Class 11"
author: "Barry Grant"
date: "5/9/2018"
output:
  word_document: default
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PDB statistics

Import our PDB statistics CSV file and calculate precent structures by experimental method.

```{r}
p <- read.csv("Data Export Summary.csv", row.names=1)
```

```{r}
percent <- (p$Total / sum(p$Total)) * 100
names(percent) <- row.names(p)
percent
```

## Using Bio3D

Load the bio3d package...

```{r}
library(bio3d)
```

Read in our HIV-Pr structure

```{r}
pdb <- read.pdb("1hsg")
pdb
```

> **Q8.** Use the Bio3D   function to write out a protein only PDB file for viewing in VMD. Also write out a second separate PDB file for the ligand with residue name MK1

```{r}
inds.ligand <- atom.select(pdb, "ligand")
inds.protein <- atom.select(pdb, "protein")
inds.protein
```

Check we have what we want

```{r}
pdb$atom[inds.ligand$atom,]
```


```{r}
pdb.protein <- trim.pdb(pdb, inds=inds.protein)
pdb.ligand <- trim.pdb(pdb, inds=inds.ligand)
pdb.ligand
```

```{r}
write.pdb(pdb.protein, file="1hsg_protein.pdb")
write.pdb(pdb.ligand, file="1hsg_ligand.pdb")

```

## Aligning multiple structures

```{r}
# Download some example PDB files
ids <- c("1TND_B","1AGR_A","1TAG_A","1GG2_A","1KJY_A","4G5Q_A")
files <- get.pdb(ids, split = TRUE)
```

```{r}
pdbs <- pdbaln(files, fit=TRUE)
```

```{r}
pdbs
```

## Search and retrieve Adenylate kinase structures

```{r}
aa <- get.seq("1ake_A")
aa
```

```{r}
# Blast or hmmer search
b <- blast.pdb(aa)
```

```{r}
hits <- plot(b)
```

```{r}
# Fetch PDBs
files <- get.pdb(hits$pdb.id, path = "pdbs", split = TRUE, gzip = TRUE)
```

---

```{r}
pdbs <- pdbaln(files, fit=TRUE)
pc <- pca(pdbs)
plot(pc)
```


View the fitted PDB

```{r, eval=FALSE}
rgl.pdbs <- view(pdbs)
rgl::rglwidget(elementId = "rgl.pdbs")
```


```{r}
# Visualize first principal component
p1 <- mktrj(pc, pc=1, file="pc_1-v2.pdb")
```

```{r, eval=FALSE}
rgl.pc1 <- view(p1)
rgl::rglwidget(elementId = "rgl.pc1")
```

---

```{r}
# Align structures
pdbs <- pdbaln(files)

# Vector containing PDB codes
ids <- basename.pdb(pdbs$id)
# Draw schematic alignment
plot(pdbs, labels=ids)
```

## Principal component analysis

```{r}
# find invariant core
core <- core.find(pdbs) 
```

```{r}
# superimpose all structures to core
pdbs$xyz = pdbfit(pdbs, core) # Perform PCA

pc.xray <- pca(pdbs)

```

```{r}
plot(pc.xray)
```


```{r}
# Visualize first principal component
pc1 <- mktrj(pc.xray, pc=1, file="pc_1.pdb")
```

# View witin R with bio3d


