---
title: "Bioinformatics Class9"
author: "Barry Grant"
date: "5/2/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Unsupervised Learning Analysis of Cancer Cells

Lets read our input data 

```{r}
url <- "https://bioboot.github.io/bggn213_S18/class-material/WisconsinCancer.csv"

# Complete the following code to input the data and store as wisc.df
wisc.df <- read.csv(url)

head(wisc.df)
```

How may diagnosis are cancer vs non cancer

```{r}
table(wisc.df$diagnosis)
```

Lets make a new data matrix with just the numeric values of intrest in it. So we want to get rid of the first 2 columns.

```{r}
# Convert the features of the data: wisc.data
# N.B.  Also remove the problem 'X' column in position 33!
wisc.data <- as.matrix( wisc.df[,-c(1:2,33)] )
rownames(wisc.data) <- wisc.df$id

#x <- c("barry", "chris", "mary", "pete")
#x[-c(1,4)]
#as.numeric(x == "mary")

diagnosis <- as.numeric(wisc.df$diagnosis == "M")
sum(diagnosis)
```

## Some questions

Explore the data you created before (wisc.data and diagnosis) to answer the following questions:

Q1. How many observations are in this dataset?
```{r}
nrow(wisc.data)
```

Q2. How many variables/features in the data are suffixed with _mean?

```{r}
#grep("_mean", colnames(wisc.data), value=TRUE, invert=TRUE)
length(grep("_mean", colnames(wisc.data)))
```

```{r}
dim(wisc.data)
```

Q3. How many of the observations have a malignant diagnosis?

```{r}
table(wisc.df$diagnosis)
```


> The functions dim(), length(), grep() and sum() may be useful for answering the first 3 questions above.



## Principal Component Analysis 

Check the mean and standard deviation of the features (i.e. columns) of the wisc.data to determine if the data should be scaled. Use the colMeans() and apply() functions like you’ve done before.

```{r}

plot( colMeans(wisc.data), type="o")

#apply(wisc.data,2,sd)
```

Lets do our PCA now

```{r}
# Perform PCA on wisc.data by completing the following code
wisc.pr <- prcomp( wisc.data, scale=TRUE )
```

How well are we doing...

```{r}
summary(wisc.pr)
```

### PCA result plots

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col=diagnosis+1)
```


## Scree-plot: variance explained.

```{r}
# Variance explained by each principal component: pve
pve <- wisc.pr$sdev^2 / sum(wisc.pr$sdev^2)

# Plot variance explained for each principal component
plot(pve, xlab = "Principal Component", 
     ylab = "Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")
```

Lets do a barplot of this data

```{r}
barplot(pve, names.arg=paste("PC",1:length(pve)), las=2, axes=FALSE, ylab="Proportion of variance" )
axis(2, at=pve, labels=round(pve,2)*100 )
```



## Hierarchical clustering

First we want to scale our data using the `scale()` function.

```{r}
# Scale the wisc.data data: data.scaled
data.scaled <- scale(wisc.data)
```

Calculate the (Euclidean) distances between all pairs of observations in the new scaled dataset and assign the result to data.dist.

```{r}
data.dist <- dist(data.scaled)
```

Create a hierarchical clustering model using complete linkage. Manually specify the method argument to hclust() and assign the results to wisc.hclust.


```{r}
wisc.hclust <- hclust(data.dist, method="complete")
```

Lets plot our tree

```{r}
plot(wisc.hclust)
abline(h=20, col="red", lwd=3)
```

### Selecting number of clusters

```{r}
wisc.hclust.clusters <- cutree(wisc.hclust, k=4)
```

### How do these groups match our `diagnosis`

```{r}
table( diagnosis )
```

```{r}
table( wisc.hclust.clusters )
```

```{r}
table( wisc.hclust.clusters, diagnosis)
```


## K-means

Input our scaled values here!

```{r}
data.scaled <- scale(wisc.data)
wisc.km <- kmeans(data.scaled, centers=2, nstart=20)
```

Cluster membership vector

```{r}
table(wisc.km$cluster)
```

Compare to expert `diagnosis`

```{r}
table(wisc.km$cluster, diagnosis)
```


# Clustering on PCA results

```{r}
## Use the distance along the first 7 PCs for clustering i.e. wisc.pr$x[, 1:7]
wisc.pr.hclust <- hclust(dist(wisc.pr$x[,1:3]), method="ward.D2")
plot(wisc.pr.hclust)
```

Lets making cluster vector

```{r}
wisc.pr.hclust.clusters <- cutree(wisc.pr.hclust, k=2)

plot(wisc.pr$x[,1:2], col=wisc.pr.hclust.clusters)
```

```{r}
table(wisc.pr.hclust.clusters, diagnosis)
```


```{r}
#install.packages("rgl")
library(rgl)

# interactive plot
#plot3d(wisc.pr$x[,1:3], type="s", col=wisc.pr.hclust.clusters)

plotids <- plot3d(wisc.pr$x[,1:3], xlab="PC 1", ylab="PC 2", zlab="PC 3", cex=1.5, size=1, type="s", col=diagnosis+1)

rglwidget(elementId = "plot3drgl")
```



## Predicting Malignancy Of New samples  

```{r}
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
```

```{r}
npc <- predict(wisc.pr, newdata=new)
```

```{r}
plot(wisc.pr$x[,1:2], col=wisc.pr.hclust.clusters)
points(npc[,1], npc[,2], col=c("purple","blue"), pch=16, cex=3)
```



