---
title: "Bioinformatics Class 12"
author: "Barry Grant"
date: "5/11/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(bio3d)
```

# Setup HIV-Pr for docking study

Get the protein first (i.e. download)
```{r}
file.name <- get.pdb("1hsg")
file.name
```

Read this file in and trim out the protein and small molecule ligand
from everything else

```{r}
hiv <- read.pdb(file.name)
hiv
```

```{r}
ligand <- trim.pdb(hiv, "ligand")
ligand
```

Extract protein 
```{r}
protein <- trim.pdb(hiv, "protein")
protein
```

```{r}
write.pdb(ligand, "1hsg_ligand.pdb")
write.pdb(protein, "1hsg_protein.pdb")
```


## Process docking result for viewing in VMD

```{r}
res <- read.pdb("all.pdbqt", multi=TRUE)
write.pdb(res, "results.pdb")
```

Compare our results to Merk Co. drug crystal structure.

```{r}
ori <- read.pdb("ligand.pdbqt")
```

```{r}
rmsd(ori, res)
```


> **Q6.** RMSD based on non hydrogen atoms

```{r}
inds <- atom.select(ori, "noh")

rmsd(ori$xyz[,inds$xyz], res$xyz[, inds$xyz])
```





## Revist Search and retrieve Adenylate kinase structures

```{r}
aa <- get.seq("P69441")
aa
```


```{r}
# Blast or hmmer search
b <- blast.pdb(aa)
plot(b)
```

```{r}
hits <- plot(b)
```

```{r}
hits$pdb.id
```

```{r}
b$hit.tbl
```

```{r}
# Side-note: reading and writting fasta
write.fasta(aa, file="ruby.fa")
aa2 <- read.fasta("ruby.fa")
aa2
```

```{r}
 # Fetch PDBs
files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```

Aling them

```{r}
pdbs <- pdbaln(files, fit=TRUE)
```


```{r}
view(pdbs)
```

```{r}
# Perform PCA & plot the results
pc.xray <- pca(pdbs) 
plot(pc.xray)
```

```{r}
pc1 <- mktrj(pc.xray, pc=1, file="pc_1.pdb")
```

```{r}
plot.pc <- view(pc1)
rgl::rglwidget(elementId = "plot.pc")
```

